---
title: "HW01-Titanic exploration"
author: "Frederike Basedow"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: github_document
---

## Load packages and data
In this homework I will explore the `Titanic` data set. This data set is included in RStudio and contains data on the survival of passengers on the Titanic. 

First I will load some packages that I would like to use:

```{r, message=FALSE}
library(tidyverse)
library(knitr)
```

Now, let's have a look at the data set!

```{r}
# Have a look at the structure of the data
str(Titanic)
kable(head(Titanic))
```

Titanic is a table, trying to get an overview with the function `head()` doesn't give as any valuable information. Let's change it to a data frame for better handling.

```{r}
# Change Titanic to data frame
Titanic_df <- data.frame(Titanic)

# Have a look at new structure
str(Titanic_df)

# Look at first few rows of new data frame
kable(head(Titanic_df))
```

Now it is easier to get an idea of the data set's contents. Let's have a closer look at the details.

```{r}
# How many rows does the data set have? This tells us how many different categories of passengers there are.
n_pass_cat <- nrow(Titanic_df)

# How many coumns does the data set have? This tells us how many different variables there are.
n_var <- ncol(Titanic_df)
```

The Titanic data set has `r n_pass_cat` different passenger categories and `r n_var` different variables.

It would be interesting to know how many passengers were on the Titanic and how many of them survived.

```{r}
# Calculate the number of passengers. We can do this by adding up all numbers in the category "Freq".
n_pass <- sum(Titanic_df$Freq)

# Calculate how many passengers survived.
n_surv <- sum(filter(Titanic_df, Survived == "Yes")$Freq)

# Calculate the percentage of survived passengers
perc_surv <- round(n_surv/n_pass*100)
```

There were `r n_pass` passengers on the Titanic and `r perc_surv`% of them survived.
